---
title: "try_rAmCharts"
author: "Ben Best"
date: "March 21, 2016"
output:
  html_document:
    css: try_rAmCharts.css
    toc: yes
  ioslides_presentation:
    incremental: true
---

## Introduction

Let's try [**rAmCharts**](https://github.com/DataKnowledge/rAmCharts), an [_htmlwidget_](http://www.htmlwidgets.org/) for [amCharts](https://www.amcharts.com/).

* [wk06_widgets](http://ucsb-bren.github.io/env-info/wk06_widgets.html#htmlwidgets)
* [Week 8: Interactive Applications - Shiny | ucsb-bren/env-info](http://ucsb-bren.github.io/env-info/wk08_shiny/)

## Methods and Results

### rAmCharts

```{r,warning=F, echo=F}
# load libaries
suppressPackageStartupMessages({
  library(pipeR)     # install.packages("pipeR")
  library(rAmCharts) # install.packages("rAmCharts")
  library(dplyr)
})

# load data
data(iris)

# generate interactive histogram
amHist(iris$Sepal.Length, freq = FALSE, breaks = 30, col = "gray",
       path = "http://www.amcharts.com/lib/3") %>% 
  plot()

# add gui images
dir.create('amcharts/images', showWarnings=F, recursive=T)
result = file.copy(
  system.file("htmlwidgets/lib/images", package = "rAmCharts"), 
  'amcharts', recursive=T)
```

### dygraphs

For time series, I recommend checking out the [dygraphs](http://rstudio.github.io/dygraphs/) library.

```{r}
# load library
library(dygraphs)

# get
lungDeaths <- cbind(mdeaths, fdeaths)

# create zoomable time-series plot
dygraph(lungDeaths) %>% 
  dyRangeSelector()
```

See also [CSS Styling for dygraphs](http://rstudio.github.io/dygraphs/gallery-css-styling.html).

## Conclusion

Whoah! This `AmCharts` JavaScript library is pretty cool, and easy to load data with `rAmCharts`.

## Discussion

- [adding CSS to rmarkdown html_document](http://rmarkdown.rstudio.com/html_document_format.html#custom_css)
- [wk06_widgets](http://ucsb-bren.github.io/env-info/wk06_widgets.html#htmlwidgets)
- [Week 8: Interactive Applications - Shiny | ucsb-bren/env-info](http://ucsb-bren.github.io/env-info/wk08_shiny/)

```{r}
help(package='rAmCharts')
```

### Leaflet 

[leaflet](http://rstudio.github.io/leaflet/) map

```{r}
library(leaflet)
library(readr)

# load data
data(quakes) # View(quakes)
quakes = quakes %>%
  mutate(row_num = row_number()) %>%
  select(row_num, everything())

# make data available
dir.create('data', showWarnings=F)
write_csv(quakes, 'data/quakes.csv')

# show `quakes` dataset, as clusters, with popup linking to row of data
leaflet(data = quakes) %>%
  addProviderTiles("Stamen.TonerLite") %>% 
  addMarkers(
    ~long, ~lat, popup = ~as.character(sprintf(
      'row_num: <a href=https://github.com/bbest/bbest.github.io/blob/master/data/quakes.csv#L%d>%s</a><br>
      mag: <strong>%s</strong><br>
      depth: <strong>%s</strong><br>', row_num+1, row_num, mag, depth)),
    clusterOptions = markerClusterOptions())
```

* [Leaflet for R - Markers](http://rstudio.github.io/leaflet/markers.html)
* [Databases with `dplyr`](https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html)

PS You can see all the code to generate this at <https://github.com/bbest/bbest.github.io/blob/master/help/try_rAmCharts.Rmd>