---
title: "wk05_ggplot"
author: "Ben Best"
date: "February 5, 2016"
output: html_document
---

## joining data

For this portion of the individual assignment, similar to last week, you'll find it easiest to copy and paste from `## joining data` onwards in  [`env-info/wk05_joining.Rmd`](https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/wk05_joining.Rmd) to your `env-info_hw/wk05_ggplot.Rmd`. Then you can play with different chunks of the code. Please be sure to answer all tasks and questions at the bottom.

The R chunks explaining the `dplyr` join functions below are pulled from the excellent [**wrangling-webinar.pdf**](wk03_dplyr/wrangling-webinar.pdf) presentation, which you should consult as you execute (see shortcuts in [rstudio-IDE-cheatsheet.pdf](refs/cheatsheets/rstudio-IDE-cheatsheet.pdf)). 

### setup

Ensure that you're in the same working directory `env-info_hw` when you Knit HTML as when you test code in the Console.

```{r setwd students}
wd = 'env-info_hw'

# set working directory for Console (vs Rmd)
if (wd %in% list.files() & interactive()){
    setwd(wd)
}

# ensure working directory
if (basename(getwd()) != wd){
  stop(sprintf("WHOAH! Your working directory is not in '%s'!\n   getwd(): %s", wd, getwd()))
}
```

### `bind_cols`

```{r bind_cols}
library(dplyr)

y = data.frame(
  x1 = c('A','B','C'), 
  x2 = c( 1 , 2 , 3), 
  stringsAsFactors=F)
z = data.frame(
  x1 = c('B','C','D'), 
  x2 = c( 2 , 3 , 4), 
  stringsAsFactors=F)
y
z
bind_cols(y, z)
```

### `bind_rows`

```{r bind_rows}
y
z
bind_rows(y, z)
```

### `union`

```{r union}
y
z
union(y, z)
```

### `intersect`

```{r intersect}
y
z
intersect(y, z)
```

### `setdiff`

```{r setdiff}
y
z
setdiff(y, z)
```

### `left_join`

```{r left_join}
songs = data.frame(
  song = c('Across the Universe','Come Together', 'Hello, Goodbye', 'Peggy Sue'),
  name = c('John','John','Paul','Buddy'), 
  stringsAsFactors=F)
artists = data.frame(
  name = c('George','John','Paul','Ringo'),
  plays = c('sitar','guitar','bass','drums'), 
  stringsAsFactors=F)
left_join(songs, artists, by='name')
```

### `inner_join`

```{r inner_join}
inner_join(songs, artists, by = "name")
```

### `semi_join`

```{r semi_join}
semi_join(songs, artists, by = "name")
```

### `anti_join`

```{r anti_join}
anti_join(songs, artists, by = "name")
```

### per capita CO<sub>2</sub> emissions 

You'll join the [population](http://data.okfn.org/data/core/population) dataset to calculate per capita CO<sub>2</sub> emissions.

_**Task**. Summarize the total emissions by country  (not World or EU28) per capita across years from your long format table and return the top 5 emitting countries for **2014**._

_**Task**. Summarize the total emissions by country (not World or EU28) per capita across years from your long format table and return the top 5 emitting countries on **average** across all years._ 

```{r read co2}
library(dplyr)
library(readxl) # install.packages('readxl')

# xls downloaded from http://edgar.jrc.ec.europa.eu/news_docs/CO2_1970-2014_dataset_of_CO2_report_2015.xls
xls = './data/co2_europa.xls'

print(getwd())
co2 = read_excel(xls, skip=12)
co2
```

```{r popn}
library(readr)
library(tidyr)

popn = read_csv(
  'https://raw.githubusercontent.com/datasets/population/master/data/population.csv')
```



---
title: "wk05_ggplot"
author: "Ben Best"
date: "February 9, 2016"
output: html_document
---

## joining data

For this portion of the individual assignment, similar to last week, you'll find it easiest to copy and paste from `## joining data` onwards in  [`env-info/wk05_joining.Rmd`](https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/wk05_joining.Rmd) to your `<user>.github.io/env-info_hw/wk05_ggplot.Rmd`. Then you can play with different chunks of the code. Please be sure to answer all tasks and questions at the bottom.

The R chunks explaining the `dplyr` join functions below are pulled from the excellent [**wrangling-webinar.pdf**](wk03_dplyr/wrangling-webinar.pdf) presentation, which you should consult as you execute (see shortcuts in [rstudio-IDE-cheatsheet.pdf](refs/cheatsheets/rstudio-IDE-cheatsheet.pdf)). 

### setup

Ensure that you're in the same working directory `env-info_hw` when you Knit HTML as when you test code in the Console.

```{r setwd students}
wd = 'env-info_hw'

# set working directory for Console (vs Rmd)
if (wd %in% list.files() & interactive()){
    setwd(wd)
}

# ensure working directory
if (basename(getwd()) != wd){
  stop(sprintf("WHOAH! Your working directory is not in '%s'!\n   getwd(): %s", wd, getwd()))
}
```

### `bind_cols`

```{r bind_cols}
library(dplyr)

y = data.frame(
  x1 = c('A','B','C'), 
  x2 = c( 1 , 2 , 3), 
  stringsAsFactors=F)
z = data.frame(
  x1 = c('B','C','D'), 
  x2 = c( 2 , 3 , 4), 
  stringsAsFactors=F)
y
z
bind_cols(y, z)
```

### `bind_rows`

```{r bind_rows}
y
z
bind_rows(y, z)
```

### `union`

```{r union}
y
z
union(y, z)
```

### `intersect`

```{r intersect}
y
z
intersect(y, z)
```

### `setdiff`

```{r setdiff}
y
z
setdiff(y, z)
```

### `left_join`

```{r left_join}
songs = data.frame(
  song = c('Across the Universe','Come Together', 'Hello, Goodbye', 'Peggy Sue'),
  name = c('John','John','Paul','Buddy'), 
  stringsAsFactors=F)
artists = data.frame(
  name = c('George','John','Paul','Ringo'),
  plays = c('sitar','guitar','bass','drums'), 
  stringsAsFactors=F)
left_join(songs, artists, by='name')
```

### `inner_join`

```{r inner_join}
inner_join(songs, artists, by = "name")
```

### `semi_join`

```{r semi_join}
semi_join(songs, artists, by = "name")
```

### `anti_join`

```{r anti_join}
anti_join(songs, artists, by = "name")
```

### per capita CO<sub>2</sub> emissions 

You'll join the [population](http://data.okfn.org/data/core/population) dataset to calculate per capita CO<sub>2</sub> emissions.

_**Task**. Calculate the per capita emissions by country (not World or EU28) and return the top 5 emitting countries for **2014**._

_**Task**. Summarize the per capita emissions by country (not World or EU28) as the mean (ie average)  value across all years and return the top 5 emitting countries._ 

```{r read co2 & popn}
library(readr)
library(readxl) # install.packages('readxl')
library(tidyr)
library(httr)   # install.packages('httr')
library(plyr)   # install.packages('plyr')
library(dplyr)

print(getwd())

# population Excel file
url = 'http://edgar.jrc.ec.europa.eu/news_docs/CO2_1970-2014_dataset_of_CO2_report_2015.xls'
xls = 'co2_europa.xls'
if (!file.exists(xls)) writeBin(content(GET(url), 'raw'), xls)

# country lookup csv file
url = 'https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/data/co2_country_to_popn.csv'
out = 'co2_country_to_popn.csv'
if (!file.exists(out)) writeBin(content(GET(url), 'raw'), out)

# read in carbon dioxide emissions file
co2 = read_excel(xls, skip=12) %>%
  rename(country=Country)

# get lookup table to translate some country names in popn to those in co2 table
cntry = read_csv(out)

# read in population and revalue some country names to match co2 table
popn = read_csv('https://raw.githubusercontent.com/datasets/population/master/data/population.csv') %>%
  select(country=`Country Name`, year=Year, popn=Value) %>%
  mutate(
    country = plyr::revalue(
      country, 
      setNames(cntry$country_co2, cntry$country_popn))) %>%
  arrange(country) %>%
  filter(year==2014) %>%
  tail(20)

```

# joining data

```{r co2 join popn & calc per capita}
co2 = read_excel(xls, skip=12) %>%
  gather(year, ktons_co2, -Country) %>%
  filter(!Country %in% c('World','EU28') & !is.na(Country)) %>%
  mutate(year = as.numeric(as.character(year))) %>%
  rename(country=Country)

co2_nopopn = co2 %>%
  filter(year == 2014) %>%
  anti_join(
    popn %>%
      filter(year==2014),
    by='country') %>%
  #arrange(desc(ktons_co2))
  arrange(country)
cat(co2_nopopn$country, sep='\n')

popn_noco2 = popn %>%
  filter(year == 2014) %>%
  anti_join(
    co2 %>%
      filter(year==2014),
    by='country') %>%
  arrange(country)
cat(popn_noco2$country, sep='\n')

d = co2 %>%
  left_join(popn, by=c('country','year')) %>%
  mutate(tons_co2_per_capita = ktons_co2 / popn * 1000)
```

_**Task**. Calculate the per capita emissions by country (not World or EU28) and return the top 5 emitting countries for **2014**._

```{r co2 per capita 2014}
d %>%
  filter(year == 2014) %>%
  arrange(desc(tons_co2_per_capita)) %>%
  head(10)
```

_**Task**. Summarize the per capita emissions by country (not World or EU28) as the mean (ie average)  value across all years and return the top 5 emitting countries._ 

```{r co2 join popn}
d %>%
  group_by(country) %>%
  summarize(
    avg_tons_co2_per_capita = mean(tons_co2_per_capita, na.rm=T)) %>%
  arrange(desc(avg_tons_co2_per_capita)) %>%
  head(5)
```


## ggplot

- <http://jcvdav.github.io/env-info_hw/wk05_ggplot.html#thin>

```{r}
library(readr)
library(ggplot2)

thindata=read_csv("./thinning.csv")

ggplot(aes(thin, cpool, group=thin), data=thindata) +
  geom_boxplot(aes(colour=thin))+
  labs(x="Thinning", y="Carbon Secuestration")+
  theme_bw()
```

